---
title: "Evaluations of positive and negative stimuli using the Affective Misattribution Procedure (AMP) and self-reports"
subtitle: "Analysis"
author: "Template: Ian Hussey; content: [Student name]"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}

# set knit options
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

# disable scientific notation
options(scipen = 999) 

```

# Dependencies

```{r}

library(tidyverse)
library(knitr)
library(kableExtra)
library(janitor)

```

# Data

Load the processed data and apply the global exclusions.

```{r}

data_processed <- read_csv("../data/processed/data_processed.csv")

data_processed_after_exclusions <- data_processed |>
  filter(exclude_participant == "include")

```

# Sample descriptives

## Sample size before exclusions

```{r}

data_processed |>
  count(name = "n") |>
  kable() |>
  add_header_above(header = c("Whole sample" = 1)) |> # note that you can add header rows to tables like this. The "1" indicates the number of columns the header should span. The sum of these numbers must equal the number of columns or you'll get an error.
  kable_classic(full_width = FALSE)

```

## Sample size after exclusions

Sample used in subsequent analyses

```{r}

data_processed_after_exclusions |>
  count(name = "n") |>
  kable() |>
  add_header_above(header = c("For analysis" = 1)) |>
  kable_classic(full_width = FALSE)

```

## Age

```{r}

data_processed_after_exclusions |>
  mutate(age = as.numeric(age)) |>
  summarise(Mean = mean(age, na.rm = TRUE),
            SD = sd(age, na.rm = TRUE)) |>
  mutate_all(.funs = janitor::round_half_up, digits = 1) |>
  kable() |>
  add_header_above(header = c("Age" = 2)) |>
  kable_classic(full_width = FALSE)

```

## Gender

```{r}

data_processed_after_exclusions |> 
  rename(Gender = gender) |>
  group_by(Gender) |> 
  summarise(n = n()) |> 
  mutate(Percent = paste0(round_half_up((n / sum(n)) * 100, 1), "%")) |>
  mutate(Gender = stringr::str_to_sentence(Gender)) |> # Change the case of the Gender variable so that it prints nicely
  kable() |>
  kable_classic(full_width = FALSE)

```

# Descriptives

Descriptive statistics and plots of the measures (excluding the demographics variables)

## Self-reported evaluations

### Descriptive stats

```{r}

# overall self-reported evaluations
dat_mean_ratings <- data_processed_after_exclusions |>
  summarise(Mean = mean(mean_evaluation, na.rm = TRUE),
            SD = sd(mean_evaluation, na.rm = TRUE)) |>
  mutate(group = "Full sample")

# self-reported evaluations by gender category
dat_mean_ratings_by_gender <- data_processed_after_exclusions |>
  group_by(group = gender) |>
  summarise(Mean = mean(mean_evaluation, na.rm = TRUE),
            SD = sd(mean_evaluation, na.rm = TRUE))

# combine both into one table
bind_rows(dat_mean_ratings,
          dat_mean_ratings_by_gender) |>
  select(Subset = group, Mean, SD) |> # select variables of interest, and rename one 
  mutate(Subset = stringr::str_to_sentence(Subset)) |> # Change the case of the Subset variable so that it prints nicely
  mutate_if(is.numeric, round_half_up, digits = 2) |>
  kable() |>
  add_header_above(header = c(" " = 1, "Self-reported evaluations" = 2)) |>
  kable_classic(full_width = FALSE)

```

### Descriptive plots

TODO add descriptive plots for self-reported evaluations

```{r}
ggplot(data_processed_after_exclusions, aes(mean_evaluation)) +
  geom_histogram(binwidth = 1,
                 fill = "white",
                 colour = "black",
                 boundary = 0.5) +
  scale_x_continuous(name = "Self Report",
                     breaks = c(1,2,3,4,5,6,7))+
  scale_y_continuous(name = "Frequency",
                     breaks = c(0,10,20,30,40,50,60,70)) +
  theme_linedraw()



```

## AMP evaluations

### Descriptive stats

TODO add descriptive stats for AMP

```{r}
# Multiple versions


# Binwidth = 0.1, starting point not one
ggplot(data_processed_after_exclusions, aes(AMP_score)) +
  geom_histogram(binwidth = 0.1,
                 fill = "white",
                 colour = "black")

# Binwidth = 0.1, starting point 0
ggplot(data_processed_after_exclusions, aes(AMP_score)) +
  geom_histogram(binwidth = 0.1,
                 fill = "white",
                 colour = "black",
                 boundary = 0.05) +
  xlab("AMP score") +
  ylab("Frequency") +
  theme_linedraw()


# seperate for gender
ggplot(data_processed_after_exclusions, aes(AMP_score, fill = gender)) +
  geom_histogram(binwidth = 0.1,
                 boundary = 0) +
  xlab("AMP score") +
  ylab("Frequency") +
  theme_linedraw()

```

### Descriptive plots

TODO add descriptive plots for AMP

```{r}


```

# Analyses & hypothesis tests

## Correlation self-reported evaluations and evaluation of the AMP

Are the two measures (which should measure the same thing) correlated?

Color the graph meaningfully (inclusive; good looking); Hexadezimalcode

```{r}
# Raw Data
ggplot(data_processed_after_exclusions, aes(x= mean_evaluation, y = AMP_score)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm")+
  theme_bw()


# Jitter 
ggplot(data_processed_after_exclusions, aes(x= mean_evaluation, y = AMP_score)) +
  geom_point(size = 2, position = "jitter") +
  geom_smooth(method = "lm")


# seperate for gender
ggplot(data_processed_after_exclusions, aes(x = mean_evaluation, y = AMP_score, fill = gender)) +
  geom_point(size = 2, position = "jitter") +
  geom_smooth(method = "lm")

```

The two measures are not correlated. The AMP is not a good measure.

## Self-reported evalautions

TODO add t test comparing self-reported evaluations between men and women

```{r}



```

## Evaluations on the Affect Misattribution Procedure

TODO add t test comparing AMP evaluations between men and women

```{r}



```

# Session info

```{r}

sessionInfo()

```
